Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-04-06 paper.qmd
di changes

# umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.8.rds")
nldr2 <- umap_pbmc |>
  ggplot(aes(x = emb1,
             y = emb2))+
  geom_point(alpha=0.1, size=1, colour=clr_choice) +
  interior_annotation("b")

nldr2c <- umap_pbmc |>
  ggplot(aes(x = emb1,
             y = emb2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("b")

@@

nldr3c <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("c")

@@

nldr4c <- tsne_pbmc |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#377eb8') +
  interior_annotation("d")

@@

nldr5c <- tsne_pbmc |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
  interior_annotation("e")

@@

nldr6c <- phate_pbmc |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("f")

@@

nldr7c <- trimap_pbmc |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("g")

@@

nldr8c <- pacmap_pbmc |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#f781bf') +
  interior_annotation("h")
```

@@

trimesh_two_curvy_removed1_tsne_with_data <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = tsne_two_curvy_scaled,
             aes(
               x = tSNE1,
               y = tSNE2
             ),
             color = clr_choice,
             alpha = 0.5,
             size = 0.5) +
  interior_annotation("b", sc_ltr_pos, cex = 1.5)
```

@@

proj_obj1_true <- proj_obj1
proj_obj1_true[["model_df"]] <- true_model_df_proj

two_curvy_proj_true <- plot_proj(
  proj_obj = proj_obj1_true, 
  point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
  plot_limits = c(-0.6, 0.6), 
  title = "a", 
  cex = 1.5, 
  axis_text_size = 4, 
  is_color = FALSE) +
  theme(aspect.ratio=1, 
        legend.position = "none")

@@

proj_obj1[["axes"]] <- axes
proj_obj1[["circle"]] <- circle

two_curvy_proj_tsne_first_model1 <- plot_proj(
    proj_obj = proj_obj1, 
    point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "c", 
    cex = 1.5, 
    position = sc_ltr_pos,
    axis_text_size = 2, 
    is_color = FALSE) +
    theme(aspect.ratio = 1,
          legend.position = "none")

@@

@@

proj_obj2_true <- proj_obj2
proj_obj2_true[["model_df"]] <- true_model_df_proj

two_curvy_proj_true2 <- plot_proj(
    proj_obj = proj_obj2_true, 
    point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
    line_param = c(0.5, 0.5, "black"), # linewidth, alpha
    plot_limits = c(-0.42, 0.42), 
    title = "b", 
    cex = 1.5, 
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

@@

proj_obj2[["axes"]] <- axes
proj_obj2[["circle"]] <- circle

two_curvy_proj_tsne_all_model1 <- plot_proj(
  proj_obj = proj_obj2, 
  point_param = c(0.1, 0.5, clr_choice), # size, alpha, color
  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
  plot_limits = c(-0.5, 0.5), 
  title = "d", 
  cex = 1, 
  position = sc_ltr_pos,
  axis_text_size = 2, 
  is_color = FALSE) #+
  #geom_segment(
  #  data = model_df, 
  #  aes(
  #    x = proj1_from, 
  #    y = proj2_from, 
  #    xend = proj1_to, 
  #    yend = proj2_to), 
  #  color = "#000000")

@@

```{r}
#| label: fig-two-curvy-true-proj
#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two non-linear clusters data (grey points) and true model (blue lines) are shown. The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape. Video of the langevitour animations are available at <>." 
#| fig-pos: H
#| fig-width: 10
#| fig-height: 5

@@

<!-- langevitour with pD model for two-curvy clusters -->
```{r}
#| label: best-fit-tsne
#| fig-cap: "Lifting the fitted model into \\pD{}. Starting from a default tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (b) and then lifted to $4\\text{-}D$ (c, d). The fit is reasonbly tight with the data, although it does not fully spread the full width. This highlights a key characteristic of tSNE, which compresses the data when applying tSNE. Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)parameter choice. Video of the langevitour animation is available at <>."
#| fig-width: 8
#| fig-height: 2

nldr_two_curvy2 + 
  trimesh_two_curvy_removed1_tsne_with_data +
  two_curvy_proj_tsne_first_model1 +
  two_curvy_proj_tsne_all_model1 +
  plot_layout(ncol = 4)
```

@@

two_curvy_proj_first_model1_umap <- plot_proj(
    proj_obj = proj_obj1, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "c2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(
        legend.position = "none"
    )

@@

proj_obj1_temp1 <- proj_obj1
proj_obj1_temp2 <- proj_obj1
proj_obj1_temp1[["model_df"]] <- model_df1

two_curvy_proj_model_delaunay_umap <- plot_proj(
    proj_obj = proj_obj1_temp1, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "a2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

proj_obj1_temp2[["model_df"]] <- model_df2

two_curvy_proj_model_benchmark2_umap <- plot_proj(
    proj_obj = proj_obj1_temp2, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "b2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 22
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

trimesh_removed_gau_tsne <- ggplot() + 
  geom_point(
    data = tsne_gau_scaled,
    aes(
      x = tSNE1,
      y = tSNE2
    ),
    colour = clr_choice,
    alpha = 0.1) + 
  geom_segment(
    data = tr_from_to_df_gau1,
    aes(
      x = x_from,
      y = y_from,
      xend = x_to,
      yend = y_to),
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  interior_annotation("a1", 
                      position = c(0.08, 0.95),
                      cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 2)
```

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 62
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

trimesh_removed_gau_umap <- ggplot() + 
  geom_point(
    data = umap_gau_scaled,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    colour = clr_choice,
    alpha = 0.05) + 
  geom_segment(
    data = tr_from_to_df_gau1,
    aes(
      x = x_from,
      y = y_from,
      xend = x_to,
      yend = y_to),
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  interior_annotation("b1", 
                      position = c(0.95, 0.95),
                      cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 2)
```

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 25
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

distance_df_small_edges_gau1 <- distance_gau1 |>
  filter(distance < benchmark_gau1) |>
  mutate(ID = row_number()) #|>
  #filter(!(ID %in% c(97)))

@@

trimesh_removed_gau_pacmap <- ggplot() + 
  geom_point(
    data = pacmap_gau_scaled,
    aes(
      x = PaCMAP1,
      y = PaCMAP2),
      colour = clr_choice,
    alpha = 0.1) +
  geom_segment(
    data = tr_from_to_df_gau1,
    aes(
      x = x_from,
      y = y_from,
      xend = x_to,
      yend = y_to),
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  interior_annotation("c1", 
                      position = c(0.08, 0.95),
                      cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_pacmap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1,
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df,
    aes(
      x = proj1_from,
      y = proj2_from,
      xend = proj1_to,
      yend = proj2_to),
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2, label=rownames(axes)),
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle,
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("c2",
                      position = c(0.08, 0.9),
                      cex = 2)

@@

error_df_one_curvy_abs <- error_df_one_curvy_abs |>
  mutate(error_cat_n = if_else(tSNE1 <= 0.25 & tSNE2 <= 0.25, "selected", "deselected")) |> ## high_error points
  mutate(error_cat_n2 = if_else(tSNE2 >= 1.2, "selected", "deselected")) ## corner points

error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs, 
      aes(x = sqrt_row_wise_total_error, 
          y = density,
          colour = factor(error_cat_n, levels = c("deselected", "selected")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("b1",
                      position = c(0.95, 0.9),
                      cex = 1.5) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))

error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs,
        aes(x = sqrt_row_wise_total_error, 
            y = density,
            colour = factor(error_cat_n2, levels = c("deselected", "selected")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("a1",
                      position = c(0.95, 0.9),
                      cex = 1.5) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))
```

```{r}
#| label: link-model-one-dens-clust
plot_tsne_dens <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n2,
                             levels = c("deselected", "selected")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  theme(aspect.ratio = 1) +
  interior_annotation("a2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

plot_tsne_dens_selected <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n,
                             levels = c("deselected", "selected")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  theme(aspect.ratio = 1) +
  interior_annotation("b2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

@@

data_c_shaped_n <- data_c_shaped |>
  select(-type) |>
  mutate(type = factor(error_df_one_curvy_abs$error_cat_n2,
                          levels=c("deselected", "selected")))

@@

langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
                         lineFrom = distance_df_small_edges_c_shaped_structure$from,
                         lineTo = distance_df_small_edges_c_shaped_structure$to,
                         group = factor(df_model_data_n$type,
                                        c("deselected", "selected", "model")),
                         levelColors = c('#d9d9d9', '#800026', "#000000"))
```

@@

proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n2,
                          levels=c("deselected", "selected"))

five_c_shaped_proj_tsne_model1 <- plot_proj(
    proj_obj = proj_obj1,
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    title = "a3", 
    cex = 1.5,
    position = c(0.08, 0.9),
    axis_text_size = 4, 
    is_color = TRUE) +
    scale_color_manual(values = c('#d9d9d9', '#800026')) +
    theme(aspect.ratio = 1,
          legend.position = "none")

proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n,
                          levels=c("deselected", "selected"))

>>>>

umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
nldr2 <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour=clr_choice) +
  interior_annotation("b")

nldr2c <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("b")

@@

nldr3c <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#f781bf') +
  interior_annotation("c")

@@

nldr4c <- tsne_pbmc |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("d")

@@

nldr5c <- tsne_pbmc |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("e")

@@

nldr6c <- phate_pbmc |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#377eb8') +
  interior_annotation("f")

@@

nldr7c <- trimap_pbmc |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
  interior_annotation("g")

@@

nldr8c <- pacmap_pbmc |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("h")
```

@@

trimesh_two_curvy_removed1_tsne_with_data <- ggplot() +
  geom_point(data = tsne_two_curvy_scaled,
             aes(
               x = tSNE1,
               y = tSNE2
             ),
             color = clr_choice,
             alpha = 0.5,
             size = 0.5) +
  geom_segment(data = tr_from_to_df_two_curvy,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  interior_annotation("a", sc_ltr_pos, cex = 1.5) +
  theme(aspect.ratio = 1)
```

@@

two_curvy_proj_true <- plot_proj(
  projected_df = projected_df, 
  model_df = true_model_df_proj, 
  axes = axes, 
  circle = circle, 
  point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
  plot_limits = c(-0.6, 0.6), 
  title = "a", 
  cex = 1.5, 
  axis_text_size = 4, 
  is_category = FALSE) +
  theme(aspect.ratio=1, 
        legend.position = "none")

@@

two_curvy_proj_tsne_first_model1 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "b", 
    cex = 1, 
    position = sc_ltr_pos,
    axis_text_size = 2, 
    is_category = FALSE) +
    theme(aspect.ratio = 1,
          legend.position = "none")

@@

projected_df <- proj_obj2$projected_df
model_df <- proj_obj2$model_df
axes <- proj_obj2$axes
circle <- proj_obj2$circle

@@

two_curvy_proj_true2 <- plot_proj(
    projected_df = projected_df, 
    model_df = true_model_df_proj, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
    line_param = c(0.5, 0.5, "black"), # linewidth, alpha
    plot_limits = c(-0.42, 0.42), 
    title = "b", 
    cex = 1.5, 
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

@@

two_curvy_proj_tsne_all_model1 <- plot_proj(
  projected_df = projected_df, 
  model_df = model_df, 
  axes = axes, 
  circle = circle, 
  point_param = c(0.1, 0.5, clr_choice), # size, alpha, color
  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
  plot_limits = c(-0.5, 0.5), 
  title = "c", 
  cex = 1, 
  position = sc_ltr_pos,
  axis_text_size = 2, 
  is_category = FALSE) +
  theme(aspect.ratio = 1)
#+
  #geom_segment(
  #  data = model_df, 
  #  aes(
  #    x = proj1_from, 
  #    y = proj2_from, 
  #    xend = proj1_to, 
  #    yend = proj2_to), 
  #  color = "#000000")

@@

```{r}
#| label: fig-two-curvy-true-proj
#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two non-linear clusters data (blue points) and true model (black lines) are shown. The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape. Video of the langevitour animations are available at <>." 
#| fig-pos: H
#| fig-width: 10
#| fig-height: 5

@@

<!-- langevitour with pD model for two-curvy clusters -->
```{r}
#| label: best-fit-tsne
#| fig-cap: "Lifting the fitted model into \\pD{}. Starting from a default tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (a) and then lifted to $4\\text{-}D$ (b, c). The fit is reasonbly tight with the data, although it does not fully spread the full width. This highlights a key characteristic of tSNE, which compresses the data when applying tSNE. Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)parameter choice. Video of the langevitour animation is available at <>."
#| fig-width: 8
#| fig-height: 2

#nldr_two_curvy2 + 
  trimesh_two_curvy_removed1_tsne_with_data +
  two_curvy_proj_tsne_first_model1 +
  two_curvy_proj_tsne_all_model1 +
  plot_layout(ncol = 3)
```

@@

two_curvy_proj_first_model1_umap <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "c2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(
        legend.position = "none"
    )

@@

two_curvy_proj_model_delaunay_umap <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df1, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "a2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

two_curvy_proj_model_benchmark2_umap <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df2, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "b2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 26
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

trimesh_removed_gau_tsne <- ggplot() + 
  geom_point(
    data = tsne_gau_scaled,
    aes(
      x = tSNE1,
      y = tSNE2
    ),
    colour = clr_choice,
    alpha = 0.1) + 
  geom_segment(
    data = tr_from_to_df_gau1,
    aes(
      x = x_from,
      y = y_from,
      xend = x_to,
      yend = y_to),
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.5) +
  interior_annotation("a1", 
                      position = c(0.08, 0.95),
                      cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.1) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 2)
```

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 90
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

trimesh_removed_gau_umap <- ggplot() + 
  geom_point(
    data = umap_gau_scaled,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    colour = clr_choice,
    alpha = 0.05) + 
  geom_segment(
    data = tr_from_to_df_gau1,
    aes(
      x = x_from,
      y = y_from,
      xend = x_to,
      yend = y_to),
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.5) +
  interior_annotation("b1", 
                      position = c(0.95, 0.95),
                      cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.1) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 2)
```

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 40
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

distance_df_small_edges_gau1 <- distance_gau1 |>
  filter(distance < benchmark_gau1) |>
  mutate(ID = row_number()) |>
  filter(!(ID %in% c(97)))

@@

trimesh_removed_gau_pacmap <- ggplot() + 
  geom_point(
    data = pacmap_gau_scaled,
    aes(
      x = PaCMAP1,
      y = PaCMAP2),
      colour = clr_choice,
    alpha = 0.1) +
  geom_segment(
    data = tr_from_to_df_gau1,
    aes(
      x = x_from,
      y = y_from,
      xend = x_to,
      yend = y_to),
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.5) +
  interior_annotation("c1", 
                      position = c(0.08, 0.95),
                      cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_pacmap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1,
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df,
    aes(
      x = proj1_from,
      y = proj2_from,
      xend = proj1_to,
      yend = proj2_to),
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.1) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2, label=rownames(axes)),
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle,
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("c2",
                      position = c(0.08, 0.9),
                      cex = 2)

@@

error_df_one_curvy_abs <- error_df_one_curvy_abs |>
  mutate(error_cat_n = if_else(tSNE1 <= 0.25 & tSNE2 <= 0.25, "high", "low")) |> ## high_error points
  mutate(error_cat_n2 = if_else(tSNE2 >= 1.2, "high", "low")) ## corner points

error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs, 
      aes(x = sqrt_row_wise_total_error, 
          y = density,
          colour = factor(error_cat_n, levels = c("low", "high")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("b1",
                      position = c(0.95, 0.9),
                      cex = 1.5) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))

error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs,
        aes(x = sqrt_row_wise_total_error, 
            y = density,
            colour = factor(error_cat_n2, levels = c("low", "high")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("a1",
                      position = c(0.95, 0.9),
                      cex = 1.5) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))
```

```{r}
#| label: link-model-one-dens-clust
plot_tsne_dens <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n2,
                             levels = c("low", "high")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  theme(aspect.ratio = 1) +
  interior_annotation("a2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

plot_tsne_dens_selected <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n,
                             levels = c("low", "high")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  theme(aspect.ratio = 1) +
  interior_annotation("b2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

@@

data_c_shaped_n <- data_c_shaped |>
  select(-type) |>
  mutate(type = factor(error_df_one_curvy_abs$error_cat_n2,
                          levels=c("low", "high")))

@@

langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
                         lineFrom = distance_df_small_edges_c_shaped_structure$from,
                         lineTo = distance_df_small_edges_c_shaped_structure$to,
                         group = factor(df_model_data_n$type,
                                        c("low", "high", "model")),
                         levelColors = c('#d9d9d9', '#800026', "#000000"))
```

@@

projected_df <- proj_obj1$projected_df
model_df <- proj_obj1$model_df
axes2 <- proj_obj1$axes
circle <- proj_obj1$circle

## To add error category
projected_df <- projected_df |>
  mutate(cluster = factor(error_df_one_curvy_abs$error_cat_n2,
                          levels=c("low", "high")))

five_c_shaped_proj_tsne_model1 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes2, 
    circle = circle, 
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    title = "a3", 
    cex = 1.5,
    position = c(0.08, 0.9),
    axis_text_size = 4, 
    is_category = TRUE) +
    scale_color_manual(values = c('#d9d9d9', '#800026')) +
    theme(aspect.ratio = 1,
          legend.position = "none")

projected_df <- projected_df |>
  select(-cluster) |>
  mutate(cluster = error_df_one_curvy_abs$error_cat_n)
